---
title: "Prior"
author: "Britta Grusdt"
date: "5/21/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(dplyr)
library(purrr)
library(ggplot2)
library(testit)
library(Metrics)
```


## Noisy independent tables

For each cell entry, a noisy value is sampled from a beta distribution parametrized
with $\alpha = \mu * v$ and $\beta = (1-\mu) * v$ where $\mu$ is the respective
cell entry. 
Plot beta distributions for each cell entry of a particular table depending on
noise parameter $v$.

```{r}
noise_v <- c(10, 50, 100, 500)
table <- c(0.7, 0.1, 0.15, 0.05)

v <- noise_v[[1]]
x <- table %>% map(function(p){
  rbeta(1000, p*v, (1-p)*v) %>% as_tibble()})  %>% bind_rows(.id = "cell") 
means <- x %>% group_by(cell) %>%  summarize(mean=mean(value))

for(v in noise_v){
  x <- table %>% map(function(p){
  rbeta(1000, p*v, (1-p)*v) %>% as_tibble()})  %>% bind_rows(.id = "cell")
  pl <- x %>% ggplot(aes(x=value, color=cell)) +
                      geom_density() +
                      facet_wrap(~cell, scales = "free_y") + 
                      labs(title =  paste("parameter v:", v)) +
                      geom_vline(aes(xintercept=mean), data=means, linetype="dotted",
                      color="red") 
  print(pl)
}
```

Load tables generated with particular beta/theta from noisy or model:
$P(C|A)=\theta + \beta * \theta - \beta, P(C|\neg A)=\beta$ and
$P(A)\sim beta(1,1)$; or if beta/theta=NA, 
$P(C|A)\sim beta(10,1), P(C|\neg A)\sim beta(1,10)$ and $P(A)\sim beta(1,1)$.

```{r}
param_theta <- NA
param_beta <- NA

target_dir <- file.path("..", "..", "data", "precomputations", "model-general",
                        sep="")

tables <- read_rds(paste(target_dir,"tables-all.rds",sep=""))
if(is.na(param_theta)){
    tables <- tables %>% filter(is.na(beta) & is.na(theta))
} else {
    tables <- tables %>% filter(beta==param_beta & theta==param_theta)
}
tables_independent <- tables %>% filter(cn=="A || C")

assert(unique(tables_independent$theta) %>% length() == 1)
assert(unique(tables_independent$beta) %>% length() == 1)
```

Compute mean/median/max/min for indpendence-score p: |P(A,C)-P(A)*P(C)| and 
also P(C|A),  the dependence-score for network A->C.

```{r }
tables_tidy <- tables %>% select(ps, vs, rowid, noise_v, cn) %>%
                unnest() %>% spread(key=vs, value=ps) %>%
               mutate(C=`AC`+`-AC`, A=`AC`+`A-C`, #p=abs(`AC`-(C*A)),
                      p_c_a=`AC` / (`AC` + `A-C`),
                      p_c_na=`-AC` / (`-AC` + `-A-C`),
                      p=abs(p_c_a - C)) %>% 
               gather(`AC`, `A-C`, `-AC`, `-A-C`, key="cell", value="val")

tables_wide <- tables_tidy %>%  spread(key=cell, value=val)

means_p <- tables_wide %>% group_by(noise_v, cn) %>%
            summarize(mean_p=mean(p, na.rm = TRUE),
                      median_p=median(p, na.rm = TRUE),
                      min_p=min(p, na.rm = TRUE), max_p=max(p, na.rm = TRUE),
                      
                      mean_p_c_a=mean(p_c_a, na.rm = TRUE),
                      median_p_c_a=median(p_c_a, na.rm = TRUE),
                      min_p_c_a=min(p_c_a, na.rm = TRUE),
                      max_p_c_a=max(p_c_a, na.rm = TRUE),
                      
                      mean_p_c_na=mean(p_c_na, na.rm = TRUE),
                      median_p_c_na=median(p_c_na, na.rm = TRUE),
                      min_p_c_na=min(p_c_na, na.rm = TRUE),
                      max_p_c_na=max(p_c_na, na.rm = TRUE),
                      )
means_p
```

Plot distributions of p (independence measure) for different noise parameters $v$.

```{r}
plot_p <- tables_wide %>% filter(cn=="A || C") %>%
            ggplot(aes(x=p,  color = noise_v)) +
              geom_density()  +
              facet_wrap(~noise_v, scales="free")
plot_p 
```

Use beta distributions (with different parameters for different noise parameter)
to approximate distributions of p (independence measure).

```{r, echo=FALSE}
noise_vals <- tables_tidy$noise_v %>% as.factor() %>%  levels() %>% as.double()
# params_beta <- c(25, 30, 35, 50, 70, 110, 150) # for p=|P(A,C)- P(A)*P(C))
# params_beta <- c(3, 3.5, 5, 7, 10, 15, 20) # for p=|P(C|A)-P(C|-A)|
params_beta <- c(8, 9, 15, 20, 25, 50, 75) # for p=|P(C|A)-P(C))|
models <- list()
i<-1
for(noise in noise_vals){
  models[[i]] <-tibble(approx=rbeta(1000, 1, params_beta[[i]]),
                       noise_v=noise, param=params_beta[[i]])
  i <- i + 1
}
beta_pdfs <- bind_rows(models)

mean_vals <- means_p %>% filter(cn=="A || C")
plot_p +
  geom_vline(aes(xintercept=mean_p), data=mean_vals, linetype="dashed") +
  geom_density(aes(x=approx, color=noise_v), data=beta_pdfs, color="green") +
  facet_wrap(~noise_v, scales="free")

```

Compute mean squared error between actual and approximated values.

```{r}
tibbles <- list()
i <- 1
tables_ind <- tables_wide %>% filter(cn=="A || C")
for(noise in noise_vals){
  data <- tables_ind %>% filter(noise_v==noise & !is.na(p)) %>% pull(p) %>% density()
  t <- tibble(x=data$x, y=data$y, noise_v=noise) %>% filter(x>=0)
  tibbles[[i]] <- t
  i <- i + 1
}
densities_tibble <- bind_rows(tibbles)

df <- densities_tibble %>% group_by(noise_v) %>% 
      mutate(f1=dbeta(x, 1, params_beta[[1]]), f2=dbeta(x, 1, params_beta[[2]]),
             f3=dbeta(x, 1, params_beta[[3]]), f4=dbeta(x, 1, params_beta[[4]]),
             f5=dbeta(x, 1, params_beta[[5]]), f6=dbeta(x, 1, params_beta[[6]]),
             f7=dbeta(x, 1, params_beta[[7]])) %>%
      gather(f1, f2, f3, f4, f5, f6, f7, key="model", value="val")

df_summary <- df %>% group_by(noise_v, model) %>% summarise(mse=mse(y,val))
df_summary 
```


Plot the distributions of p (independence measure) and of P(C|A) (dependence measure
for network A->C) for all networks.


```{r}
noise <- 100
mean_vals <- means_p %>% filter(noise_v==noise)

tables_wide %>% filter(noise_v==noise) %>% 
  ggplot() + geom_density(aes(x=p,color=cn)) + facet_wrap(~cn, scales="free") +
  geom_vline(aes(xintercept=mean_p), data=mean_vals, linetype="dashed") +
  geom_vline(aes(xintercept=median_p), data=mean_vals, linetype="dotted",
             color="red") +
  # labs(title = "p=|P(A,C) - P(A) * P(C)|")
  labs(title = "p=|P(C|A) - P(C|-A)|")

tables_wide %>% filter(noise_v==noise) %>% 
  ggplot() + geom_density(aes(x=p_c_a, color=cn)) +
  facet_wrap(~cn, scales="free") +
  geom_vline(aes(xintercept=mean_p_c_a), data=mean_vals, linetype="dashed") +
  geom_vline(aes(xintercept=median_p_c_a), data=mean_vals, linetype="dotted",
             color="red") +
  labs(title = "P(C|A)")


tables_wide %>% filter(noise_v==noise) %>% 
  ggplot() + geom_density(aes(x=p_c_na, color=cn)) +
  facet_wrap(~cn, scales="free") +
  geom_vline(aes(xintercept=mean_p_c_na), data=mean_vals, linetype="dashed") +
  geom_vline(aes(xintercept=median_p_c_na), data=mean_vals, linetype="dotted",
             color="red") +
  labs(title = "P(C|-A)")

```



